Notes of Basic Topolgy

### 2016-11-19 Basic Topology Notes/Basic Topology.tex
Finished chapter 4.4, orbit space of Basic Topology.

A bit more examples about matrices:
\begin{ex}
    \nomen{$\mathbb{M}(n)$} the $n\times n$ matrices, is not a
    topological group. But its subspace
    \nomen{$\mathrm{GL(n)}$}, specifically,
    $\mathrm{GL(n,\mathbb{R})}$ or $\mathrm{GL(n,\mathbb{C})}$, is a
    topological group. This is demonstrated in page 76, theorem 4.12.
\end{ex}
\begin{fact}
    $\mathrm{GL(n)}$ is not compact. It has two idsjoint nonempty open
    sets: those with positive and those with negative determinants.
\end{fact}
\begin{thm}
    $\mathrm{O}(n)$ and $\mathrm{SO}(n)$ are closed and compact.
    $\mathrm{SO}(n)$ is a sub-$\mathcal{TPG}$ of $\mathrm{O}(n)$.
\end{thm}
\begin{fact}
    $\mathrm{SO}(2)\cong S^1$ and $\mathrm{SO}(3)\cong P^3$.
\end{fact}
\begin{remark}
    These two facts established on page 77. The first one can be eaily
    guess. Since a rotation is obviously determined by a rotation
    degree on $S^1$. Mathematically we have
    \begin{equation}
        \left(\begin{array} {cc}
            \cos\theta & -\sin\theta \\
            \sin\theta & \cos\theta 
        \end{array}\right) \cong e^{i\theta}
    \end{equation}

    The second one has a physical picture. The $P^3$ lies in
    $\mathbb{H}$ (TBD). % TODO understand this. Related: Geometric algebra
    % https://en.wikipedia.org/wiki/Geometric_algebra 
\end{remark}

\section{Anchor}
\label{sec:Anchor}

>>>>

A bit more examples about matrices:
\begin{ex}
    \nomen{$\mathbb{M}(n)$} the $n\times n$ matrices, is not a
    topological group. But its subspace
    \nomen{$\mathrm{GL(n)}$}, specifically,
    $\mathrm{GL(n,\mathbb{R})}$ or $\mathrm{GL(n,\mathbb{C})}$, is a
    topological group. This is demonstrated in page 76, theorem 4.12.
\end{ex}
\begin{fact}
    $\mathrm{GL(n)}$ is not compact. It has two idsjoint nonempty open
    sets: those with positive and those with negative determinants.
\end{fact}
\begin{thm}
    $\mathrm{O}(n)$ and $\mathrm{SO}(n)$ are closed and compact.
    $\mathrm{SO}(n)$ is a sub-$\mathcal{TPG}$ of $\mathrm{O}(n)$.
\end{thm}
\begin{fact}
    $\mathrm{SO}(2)\cong S^1$ and $\mathrm{SO}(3)\cong P^3$.
    Here $\cong$ means isomorphisms of topological groups.
\end{fact}
\begin{remark}
    These two facts established on page 77. The first one can be eaily
    guess. Since a rotation is obviously determined by a rotation
    degree on $S^1$. Mathematically we have
    \begin{equation}
        \left(\begin{array} {cc}
            \cos\theta & -\sin\theta \\
            \sin\theta & \cos\theta 
        \end{array}\right) \cong e^{i\theta}
    \end{equation}

    The second one is proved mathematical in book \cite{book}. But it
    has a physical argument. Remember we have the the homogeneous
    coordinates for $P^3$, such as $[1,\theta_x,\theta_y,\theta_z]$.
    As indicated in my labels, the three free coordinates $\theta_i$
    can be regarded as rotation in 3-dimensional space. This rotation
    preserves the orientation, so it is in $\mathrm{SO}$, not in
    $\mathrm{O}$. 
\end{remark}

\subsection{Orbit Space}
\label{sec:Orbit-Space}
\begin{defi}[Group Action on Topology Space]
\nomenclature{Group Action on Topology Space}{\nomrefpage.}
    A topological group $G$ is said to act as a group of
    homeomorphisms on a space $X$ if each group element (let $g,h\in G$)
    induces a homeomorphism of the space in such a way that:
    \begin{enumerate}
        \item $(hg)(x) = h(g(x)),\, \forall x\in X$;
        \item $e(x)=x,\, \forall x\in X$, where $e=gg^{-1}$;
        \item the function $G\times X\to X$,$(g,x)\mapsto g(x)$ is
            continusous.
    \end{enumerate}
\end{defi}

The subset of $X$, consisting of $g(x)$ for all $g\in G$, is called an
\nomen{orbit} of $x\in X$, written $O(x)$. Thought, it more convenient to
write it just as \nomen{$Gx$}, as in textbooks of abstract algebra.

\begin{fact}
    A common fact in abstract algebra here is: each orbit $Gx$ is
    disjoint. If two $Gx\cap Gy\neq \varnothing$, then $Gx=Gy$.
\end{fact}

By above fact, orbits partitions $X$, hence we can form the
Identification space, with every elements in $X$ identified with their
brothers in the same orbit. The result is \nomen{orbit space $X/G$}.

\begin{ex}
    \label{ex:R-over-Z-T}
    $\mathbb{Z}$ acts on $\mathbb{R}$ by addition $x\mapsto x+n$,
    $x\in \mathbb{R},\,n\in\mathbb{Z}$. It partitioned $\mathbb{R}$
    into intervals, for each $x\in X$, $x\sim x+n,\forall n\in
    \mathbb{Z}$. The orbit space $\mathbb{R}/\mathbb{Z}$ is
    homeomorphic to $S^1$.
\end{ex}

An action $G$ on $X$ is called \nomen{transitive}, if and only if the
orbit space $X/G$ is the trivial point $\{1\}$. Or equivalently, the
only orbit is the whole space, i.e. $Gx=G,\,\forall x\in G$.

\begin{ex}
    The orthogonal action $O(n)$ on $S^{n-1}$ is transitive.
    Physically, this is saying that $\forall x\in S^{n-1}$, it can be
    rotated into $\forall y\in S^{n-1}$. A mathematical proof is on
    page 79 of \cite{book}
\end{ex}

\paragraph{A lot of examples from book \cite{book}}$ $

\begin{ex}
    Extending example \ref{ex:R-over-Z-T}:
\begin{equation}
    \mathbb{E}^2/(\mathbb{Z}\times\mathbb{Z}) = T\text{ (torus)}
\end{equation}
    Here $=$ means homeomorphism.
\end{ex}
\begin{ex}
    \begin{equation}
        S^n/\mathbb{Z}_2 = P^n
    \end{equation}
    Here $=$ means homeomorphism.
\end{ex}
\begin{ex}[Three ways of $\mathbb{Z}_2$ acting on $T$]
    The detailed procedure is to be found on page 91 of \cite{book}.
    Here's a picture to visualize the action:
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\linewidth]{pics/3-ways-Z2-on-T.PNG}
        \caption{$ $}
    \end{figure}
    The results are (a) a sphere; (b) a torus; (c) a Klein bottle.
\end{ex}
\begin{ex}
    If $G$ is a topological group, and $H$ is
    $\mathcal{TPG}$-subgroup. Then, the left cosets of right cosets
    can be canonically seen as orbits. See more on page 81, example 4.
\end{ex}
\begin{ex}
\begin{align}
    \mathrm{O}(n)/\mathrm{O}(n-1) &= S^{n-1} \\
    \mathrm{SO}(n)/\mathrm{SO}(n-1) &= S^{n-1}
\end{align}
    Here $=$ means homeomorphism. The first is established
    mathematically in page 82 of \cite{book}. The second is mentioned
    there, indicating a similar proof.

    Here I give an argument. Consider a unit vector $y$ in $S^{n-1}$, if
    we want to rotate another unit vector $e_1$ to $y$, since the
    action is transitive, we can easily find a $A\in \mathrm{O}(n)$ to
    do this. But in addition, we can also find that $A\cdot B$, where
    $B\in \mathrm{O}(n-1)$ rotates the space around $e_1$ (thus
    leaving $e_1$ un-affected) also do our job. So there is an
    $\mathrm{O}(n-1)$ redundancy in $\mathrm{O}(n)\to S^{n-1}$.
    Similar for the second relation.
\end{ex}

\begin{thm}
    Let $G$ acts on $X$ and suppose that both $G$ and $X/G$ are
    connected, then $X$ is connected.
\end{thm}
\begin{fact}
    Using the theorem above, one can deduce that: $\mathrm{SO}(1)$ is
    connected, $S^{n-1}$ is connected, so $\mathrm{SO}(n)$ is
    connected.
\end{fact}

Next, the book \cite{book} (page 82 to 85) introduces several three
spaces (\nomen{Lens space}, \nomen{irrational flow} on $T$ torus,
\nomen{fundamental region} or in my word \textit{space filling
shapes}) and two group \nomen{Euclidean group} (page 84) and
\nomen{plane-crystallographic group} (page 85). To save time, I leave
here only some pictures:
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/irrational-flow.png}
    \caption{Irrational Flow on $T$}
\end{figure}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{pics/space-filling-shapes.png}
    \caption{Space-filling Shapes}
\end{figure}
\section{Anchor}
\label{sec:Anchor}
